---
title: "Data + Desserts"
output: github_document
---

### What does our data look like?
```{r}
library(YaleToolkit)
df <- read.csv("sales.csv", stringsAsFactors = FALSE,
               header = TRUE, blank.lines.skip = FALSE, as.is = TRUE)

whatis(df) ## high level view of our data

```

A really powerful tool I like to use is sapply. Here we'll use it to create a frequency table of each column of the dataframe, and get a view for what our data looks like. 

```{r}
## excluding order_id and ordered_at to reduce noise
sapply(names(df[,-c(1,3)]), function(x) table(df[x])) ## frequency tables of each column
# sapply(names(df), function(x) table(df[x])) ## frequency tables of each column
```
We see a few problems from the getgo:

* order_id is not unique
* name includes product and flavors
* price and line_total include "-" for prices
* name includes empty strings

Each section to follow will clean up what we found!

### Removing Duplicate Rows
```{r}
df <- unique(df)
```

### Removing rows with no entries in name column
```{r}
no_name <- sapply(c(1:nrow(df)), ## identifying names that are empty strings
                  function(x) nchar(df[x, "name"]) == 0) 
df <- df[no_name == FALSE,] ## removed rows with no name
```

### Cleaning up price and line_total 
* negative values
* recreating line_total column
```{r}
df$price <- gsub("-", "", df$price)  ## removing - signs from prices
df$price.f <- gsub("\\$", "", df$price) ## removing $ sign from prices
df$price.f <- as.numeric(df$price.f)  ## turning char into numeric
df$line_total <- df$price.f * df$quantity  ## recalculating line_total
df$line_total.f <- df$line_total  ## creating column with float values
df$line_total <- paste("$", df$line_total, sep = "")  ## returning to char
## gets prices with change ie. $10.5 and adds a 0 to the end
df$line_total[c(grep("\\.", df$line_total))] <- paste(df$line_total[c(grep("\\.",
                                                      df$line_total))], "0", sep = "")
## this gets the prices that dont have change ie. $3, and replaces with $3.00
df$line_total[-c(grep("\\.", df$line_total))] <- paste(df$line_total[-c(grep("\\.",
                                                df$line_total))], ".00", sep = "")  ## adding .00 to end of prices
head(df)
```

### Making unique order_id's
```{r}
# order_id is not unique
df$order_id <- c(1:nrow(df))  ## replacing with unique numbers [1:27903]
length(unique(df$order_id))
```

### Cleaning up name column
* product column
* name (flavor) column
```{r, warning=FALSE}
x <- data.frame(t(sapply(df$name, function(y) strsplit(y,split="\" ")[[1]])))
names(x) <- c("product", "name")  ## renaming col names
x$product <- gsub("\"", "", x$product) ## removing any quotes from products 
df <- df[,-which(names(df) %in% "name")]  ## removing old name column (programatically)
df <- cbind(df, x)  ## adding new cols to og df
df$name <- as.character(df$name)
head(df)
#sapply(names(df), function(x) table(df[x])) ## frequency tables of each column 
```

### Barplots

```{r}
#### icecream profits barplot ordered by greatest to least $$
p <- tapply(df$line_total.f, df$name, sum) ## total profit by flavor
p <- p[order(p, decreasing = TRUE)]  ## ordering least to greatest
barplot(p, main = "Ice Cream Profits", las = 2, col = rainbow(length(names(table(df$name)))))
```
```{r}
## product profits ordered by greatest to least $$
c <- tapply(df$line_total.f, df$product, sum) ## total profit by flavor
c <- c[order(c, decreasing = TRUE)]  ## ordering least to greatest
barplot(c, main = "Product Profits", las = 2, col = rainbow(length(names(table(df$product)))))
```